full title

### 2020-11-21 paper/pldi21/ds.tex
Till Sec. 3.5

\subsection{Combined Domain}

In this part, we define a \textit{combined domain} of a given sensitive abstract
domain with the sealed symbolic domain and its one-step execution.
\begin{definition}[Combined Domain]
  A \textit{combined domain} is $\combdom = \sabsdom \times \symbdom$ and its
  concretization function $\combgamma: \combdom \rightarrow \dom$ and join
  operator defined as
  \begin{align}
    \combgamma((\sabselem, \symbelem)) &= \sgamma(\sabselem) \cup
      \symbgamma(\symbelem)\\
    (\sabselem, \symbelem) \join ({\sabselem}', \symbelem') &= (\sabselem \join
      {\sabselem}', \symbelem \cup \symbelem')
  \end{align}
\end{definition}

Before defining the one-step execution for the combined domain, we introduce
\textit{analysis elements} to easily configure different types of abstract
states in sensitive abstract domain and sealed symbolic domain.
\begin{definition}[Analysis Elements]\label{def:aelem}
  An \textit{analysis element} $\aelem \in \aelemset = (\viewset \times \absdom)
  \uplus (\absimapset \times \symbstset)$ is either 1) a pair of a view and an
  abstract state in sensitive abstract domain $\sabsdom$, or 2) a pair of an
  abstract instantiation map and a sealed symbolic state in sealed symbolic
  domain $\symbdom$.  Its concretization function $\aelemgamma:
  \aelemset \rightarrow \dom$ is defined as
  \[
    \aelemgamma(\aelem) = \left\{
      \begin{array}{ll}
        \viewmap(\view) \cap \gamma(\abselem) & \text{if} \; (\view, \abselem) = \aelem\\
        \instant{\symbst}{\absimap} & \text{if} \; (\absimap, \symbst) = \aelem\\
      \end{array}
    \right.
  \]
\end{definition}

@@

Moreover, to freely convert types of analysis elements, we define two converters:
\begin{align}
  \asconverter & : (\viewset \times \absdom) \hookrightarrow
    (\absimapset \times \symbstset)\\
  \saconverter & : (\viewset \times
    \absdom) \leftarrow (\absimapset \times \symbstset)
\end{align}
While the converter $\saconverter$ is total, the other one $\asconverter$ is
\textit{partial}. Thus, it is possible to convert an analysis element
$(\view, \abselem)$ in sensitive abstract domain to another analysis element in
sealed symbolic domain only if the convert $\asconverter$ is defined: $(\view,
\abselem) \in \Dom(\asconverter)$.  In addition, they should convert given
elements without loss of information for all $\aelem \in \aelemset$:
\[
  \asconverter(\aelem) = \aelem' \Rightarrow \left\{
  \begin{array}{l}
    \aelem = \saconverter(\aelem')\\
    \aelemgamma(\aelem) = \aelemgamma(\aelem')\\
  \end{array}
  \right.
\]

Now, we define the \textit{combined one-step execution} $\combstep: \combdom
\rightarrow \combdom$ with two converters $\asconverter$ and $\saconverter$ and
it consists of two steps; 1) the \textit{triage} step prepares the execution of
analysis elements by converting their types via two converters according to
their status, and 2) the \textit{execution} step performs execution of each
analysis element using abstract one-step execution $\absstep$ in sensitive
abstract domain and sealed symbolic one-step execution $\symbstep$ in sealed
symbolic execution.
\begin{definition}[Combined One-Step Execution]
  A \textit{combined one-step execution} $\combstep: \combdom \rightarrow
  \combdom$ is define as
  \[
    \combstep(\combelem) = (\sabsstep(\sabselem), \symbstep(\symbelem))
  \]
  where $(\sabselem, \symbelem) = \triage(\combelem)$.
\end{definition}

The $\triage$ function decides whether a new sealed symbolic execution is
\textit{introduced} or an existing one is \textit{terminated}.  When an analysis
element $(\view, \abselem)$ in the sensitive abstract domain is defined in the
converter $\asconverter$, a new sealed symbolic execution is introduced by
converting the analysis element to the corresponding one $(\absimap, \symbst) =
\asconverter((\view, \abselem))$ in the sealed symbolic domain.  In the other
hand, assume that if there is no more other sealed symbolic states having
symbolic transitions with $\symbst$ for an analysis element $(\absimap,
\symbst)$ in the sealed symbolic domain; $\symbst \symbtrans \excst$.  Then, the
sealed symbolic execution for $(\absimap, \symbst)$ is terminated by 1)
converting it to another analysis element $(\view, \abselem) =
\saconverter((\absimap, \symbst))$ in the sensitive abstract domain and 2)
merging with the current abstract state stored in the view $\view$ with
$\abselem$.

To formally define the $\triage$ function, we first define a $\atriage$ function
for analysis elements using two converters and then define the $\triage$ function.
\begin{definition}[$\atriage$]\label{def:atriage}
  The function $\atriage: \aelemset \rightarrow \aelemset$ for analysis elements
  is defined as
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \aelem \in
        \Dom(\asconverter)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = (\absimap, \symbst) \wedge \symbst \symbtrans
        \bot\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
\end{definition}
\begin{definition}[$\triage$]\label{def:triage}
  The triage function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as
  \[
    \triage((\sabselem, \symbelem)) = \left(
      \lambda \view. \bigjoin \{ \abselem \!\mid\! (\view, \abselem) \in E \},
      E \cap (\absimapset \times \symbstset)
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \sabselem(\view)) \mid \view \in \viewset \} \cup \symbelem)
  \]
  (The dot notation $\dot{f}$ denotes the element-wise extended function of a
  function $f$.)
\end{definition}

@@

Figure~\ref{fig:ds-examples}(b) shows the analysis with the combined domain when
the initial value of $\varx$ is $0$.  First, in the triage step,
the converter $\asconverter$ converts the analysis element $(\lab_0, 0)$ to
another analysis element $\langle \varnothing, 0 \rangle$ with the label
$\lab_0$.  Moreover, it does not introduce any sealed symbolic values because
the value represents only a single value.  Until the end of the program, the
sealed symbolic execution from $\langle \varnothing, 0 \rangle$ is successfully
continues.  However, there is no more possible symbolic transition for the
symbolic state $\langle \varnothing, 0 \rangle$ with the label $\lab_4$.  Thus,
it is converted to $(\lab_4, 0)$ via the converter $\saconverter$.

>>>>

\subsection{Combined Domain}
We now define a \textit{combined domain} of a given sensitive abstract
domain with the sealed symbolic domain and its one-step execution.
\begin{definition}[Combined Domain]
  A \textit{combined domain} is $\combdom = \sabsdom \times \symbdom$ and its
  concretization function $\combgamma: \combdom \rightarrow \dom$ and join
  operator are defined as follows:
  \begin{align}
    \combgamma((\sabselem, \symbelem)) &= \sgamma(\sabselem) \cup
      \symbgamma(\symbelem)\\
    (\sabselem, \symbelem) \join ({\sabselem}', \symbelem') &= (\sabselem \join
      {\sabselem}', \symbelem \cup \symbelem')
  \end{align}
\end{definition}

Before defining the one-step execution for the combined domain, we introduce
\textit{analysis elements} to easily configure different types of abstract
states in the sensitive abstract domain and the sealed symbolic domain.
\begin{definition}[Analysis Elements]\label{def:aelem}
  An \textit{analysis element} $\aelem \in \aelemset = (\viewset \times \absdom)
  \uplus (\absimapset \times \symbstset)$ is either 1) a pair of a view and an
  abstract state in a sensitive abstract domain $\sabsdom$, or 2) a pair of an
  abstract instantiation map and a sealed symbolic state in a sealed symbolic
  domain $\symbdom$.  Its concretization function $\aelemgamma:
  \aelemset \rightarrow \dom$ is defined as follows:
  \[
    \aelemgamma(\aelem) = \left\{
      \begin{array}{ll}
        \viewmap(\view) \cap \gamma(\abselem) & \text{if} \; (\view, \abselem) = \aelem\\
        \instant{\symbst}{\absimap} & \text{if} \; (\absimap, \symbst) = \aelem\\
      \end{array}
    \right.
  \]
\end{definition}

@@

Moreover, to freely convert between different kinds of analysis elements, we define two converters:
\begin{align}
  \asconverter & : (\viewset \times \absdom) \hookrightarrow
    (\absimapset \times \symbstset)\\
  \saconverter & : (\viewset \times
    \absdom) \leftarrow (\absimapset \times \symbstset)
\end{align}
While the converter $\saconverter$ is total, the other one $\asconverter$ is
\textit{partial}. Thus, it is possible to convert an analysis element
$(\view, \abselem)$ in a sensitive abstract domain to another analysis element in
a sealed symbolic domain only if the convert $\asconverter$ is defined: $(\view,
\abselem) \in \Dom(\asconverter)$.  In addition, they should convert given
analysis elements without loss of information for all $\aelem \in \aelemset$:
\[
  \asconverter(\aelem) = \aelem' \Rightarrow \left\{
  \begin{array}{l}
    \aelem = \saconverter(\aelem')\\
    \aelemgamma(\aelem) = \aelemgamma(\aelem')\\
  \end{array}
  \right.
\]

Now, we define the \textit{combined one-step execution} $\combstep: \combdom
\rightarrow \combdom$ with two converters $\asconverter$ and $\saconverter$.
It consists of two steps: 1) the \textit{\triagename} step converts
analysis elements if a new sealed symbolic execution starts or an
existing one stops, and 2) the \textit{execution} step performs execution of each
analysis element using the abstract one-step execution $\absstep$ in the sensitive
abstract domain and the sealed symbolic one-step execution $\symbstep$ in the sealed
symbolic execution.
\begin{definition}[Combined One-Step Execution]
A \textit{combined one-step execution} $\combstep: \combdom \rightarrow
\combdom$ is define as follows:
  \[
    \combstep(\combelem) = (\sabsstep(\sabselem), \symbstep(\symbelem))
  \]
where $(\sabselem, \symbelem) = \triage(\combelem)$.
\end{definition}

From a given $\combelem$, the $\triage$ function makes analysis elements
and converts them if a new sealed symbolic execution
begins or an existing sealed symbolic execution terminates.
Specifically, for an analysis element $(\view, \abselem)$ in the sensitive abstract domain,
if the converter $\asconverter$ is defined for it, $\triage$ introduces a new sealed symbolic execution
by converting the analysis element to its corresponding one $(\absimap, \symbst) =
\asconverter((\view, \abselem))$ in the sealed symbolic domain.
On the other hand, for an analysis element $(\absimap, \symbst)$ in the sealed symbolic domain,
if it does not have any sealed symbolic states to transit to, $\symbst \symbtrans \excst$,
the sealed symbolic execution for $(\absimap, \symbst)$ terminates;
It converts the analysis element to its corresponding one $(\view, \abselem) =
\saconverter((\absimap, \symbst))$ in the sensitive abstract domain and
merges the current abstract state stored in the view $\view$ with $\abselem$.

To formally define the $\triage$ function, we first define a $\atriage$ function
for analysis elements using two converters.
\begin{definition}[$\atriage$]\label{def:atriage}
  The function $\atriage: \aelemset \rightarrow \aelemset$ for analysis elements
  is defined as follows:
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \aelem \in
        \Dom(\asconverter)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = (\absimap, \symbst) \wedge \symbst \symbtrans
        \bot\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
\end{definition}
\begin{definition}[$\triage$]\label{def:triage}
  The \triagename function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as follows:
  \[
    \triage((\sabselem, \symbelem)) = \left(
      \lambda \view. \bigjoin \{ \abselem \!\mid\! (\view, \abselem) \in E \},
      E \cap (\absimapset \times \symbstset)
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \sabselem(\view)) \mid \view \in \viewset \} \cup \symbelem)
  \]
and the dot notation $\dot{f}$ denotes the element-wise extended function of a
function $f$.
\end{definition}

@@

Figure~\ref{fig:ds-examples}(b) shows the analysis with the combined domain when
the initial value of $\varx$ is $0$.  First, in the \triagename step,
the converter $\asconverter$ converts the analysis element $(\lab_0, 0)$ to
another analysis element $\langle \varnothing, 0 \rangle$ with the label
$\lab_0$.  Moreover, it does not introduce any sealed symbolic values because
the value represents only a single value.  Until the end of the program, the
sealed symbolic execution from $\langle \varnothing, 0 \rangle$ is successfully
continues.  However, there is no more possible symbolic transition for the
symbolic state $\langle \varnothing, 0 \rangle$ with the label $\lab_4$.  Thus,
it is converted to $(\lab_4, 0)$ via the converter $\saconverter$.
